---
title: "Project Component 1"
author: "Himanshu, MDS202327"
date: "`r Sys.Date()`"
output:
#   bookdown::pdf_document2:
#     latex_engine: xelatex
# urlcolor: blue
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Introduction

In this report we try to discover patterns and make inferences about the pollution level in stations in and around Delhi. We also look at the relation between the pollutant parameters and try to find which parameter are related to each other and see if there is a pattern across all ten stations.

## Data Description

The data contains 10,600 rows and 9 columns, namely, siteName, siteCode, Date and six air pollution parameters i.e. PM2.5, PM10, NO\textsubscript{2}, NH\textsubscript{3}, SO\textsubscript{2}, Ozone for ten stations in New Delhi, collected from CPCB website from 08-02-2018 to 02-01-2021 on daily basis. There are 1060 entries for each station, one for all the dates between 08-02-2018 and 02-01-2021 (both inclusive). The data for the parameters is average of 24 hour data collected every 15 minutes. The units for all the parameters in the data are $\frac{\text{ug}}{\text{m}^3}$ that represents micrograms(one-millionth of a gram) of a gaseous pollutant per cubic meter of air.

```{=tex}
\begin{table}[h!]
\centering
%\renewcommand{\arraystretch}{1.35}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
  \hline
  \textbf{siteName}&\textbf{siteCode}&\textbf{Date}&\textbf{PM2.5}&\textbf{PM10}&\textbf{NO\textsubscript{2}}&\textbf{NH\textsubscript{3}}&\textbf{SO\textsubscript{2}}&\textbf{Ozone}\\
  {\scriptsize<chr>}&{\scriptsize<int>}&{\scriptsize<chr>}&{\scriptsize<dbl>}&{\scriptsize<dbl>}&{\scriptsize<dbl>}&{\scriptsize<dbl>}&{\scriptsize<dbl>}&{\scriptsize<dbl>}\\ \hline
  Sonia Vihar&1432&2019-09-19&17.62&65.71&13.18&26.37&12.64&36.09\\[1.05ex] \hline
  Jahangirpuri&1423&2020-03-01&51.20&120.17&72.40&36.34&2.04&12.23\\[1.05ex] \hline
  Wazirpur&1434&2020-04-12&44.46&85.50&32.24&23.36&14.07&52.15\\[1.05ex] \hline      
  Najafgarh&1427&2018-05-19&100.06&287.78&28.60&46.65&7.63&73.52\\[1.05ex] \hline
  Patparganj&1431&2018-10-27&189.89&384.89&63.65&85.26&4.39&18.85\\[1.05ex] \hline
\end{tabular}
\caption{A glimpse of random sample of the data.}
\label{table1}
\end{table}
```

The names of all ten stations with their respective site codes are displayed in the table below.

```{=tex}
\begin{table}[h!]
\centering
%\renewcommand{\arraystretch}{1.35}
\begin{tabular}{|c|c|c|c|c|c|}
  \hline
  \textbf{Site Name}&Ashok Vihar&Dwarka-Sector&Jahangirpuri&Najafgarh&Narela\\[1.05ex] \hline
  \textbf{Site Code}&1420&1422&1423&1427&1426\\ \hline\hline
  \textbf{Site Name}&Patparganj&Rohini&Sonia Vihar&Vivek Vihar&Wazirpur\\[1.05ex] \hline
  \textbf{Site Code}&1431&1430&1432&1435&1434\\ \hline 
\end{tabular}
\caption{Site Names and corresponding Site Codes}
\label{table2}
\end{table}
```
\newpage

## Exploratory Data Analysis

<!-- We use the following libraries for handling data and creating plots for this report. -->

```{r, results='hide', include=FALSE}
# Importing libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(knitr)
library(reshape2)
```

```{r, include=FALSE}
# We read our data into a data frame
# library(bookdown)
df <- read.csv("delhi.csv", header = TRUE)
```

Since our data has a date column, we would want to exploit it to our use to plot some time-series plots and analysis. As displayed in table \ref{table1}, the data column has type chr, so we must first convert it to date type.

```{r,results='hide'}
#Changing Date data type from chr to date
df$Date <- as.Date(df$Date)
```

```{=tex}
\begin{table}[h!]
\centering
%\renewcommand{\arraystretch}{1.35}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
  \hline
  \textbf{siteName}&\textbf{siteCode}&\textbf{Date}&\textbf{PM2.5}&\textbf{PM10}&\textbf{NO\textsubscript{2}}&\textbf{NH\textsubscript{3}}&\textbf{SO\textsubscript{2}}&\textbf{Ozone}\\
  {\scriptsize<chr>}&{\scriptsize<int>}&{\scriptsize<date>}&{\scriptsize<dbl>}&{\scriptsize<dbl>}&{\scriptsize<dbl>}&{\scriptsize<dbl>}&{\scriptsize<dbl>}&{\scriptsize<dbl>}\\ \hline
  Sonia Vihar&1432&2019-09-19&17.62&65.71&13.18&26.37&12.64&36.09\\[1.05ex] \hline
  Jahangirpuri&1423&2020-03-01&51.20&120.17&72.40&36.34&2.04&12.23\\[1.05ex] \hline
  Wazirpur&1434&2020-04-12&44.46&85.50&32.24&23.36&14.07&52.15\\[1.05ex] \hline      
  Najafgarh&1427&2018-05-19&100.06&287.78&28.60&46.65&7.63&73.52\\[1.05ex] \hline
  Patparganj&1431&2018-10-27&189.89&384.89&63.65&85.26&4.39&18.85\\[1.05ex] \hline
\end{tabular}
\caption{A glimpse of random sample of the data after chnaging type of Date column.}
\label{table3}
\end{table}
```

\vspace{1cm}

```{=tex}
\flushleft
\textbf{\large An Histogram}
```

We would naturally want to see the range of the values attained by all the parameters and the density with which these are attained. Hence, histograms come handy.

```{r, echo=FALSE,  fig.width=10, fig.height=6}
par(mfrow = c(2,3))
hist(df$PM2.5, probability = TRUE, main = "", xlab = "PM2.5", ylab = "")
hist(df$PM10, probability = TRUE, main = "", xlab = "PM10", ylab = "")
hist(df$NO2, probability = TRUE, main = "", xlab = "NO2", ylab = "")
hist(df$NH3, probability = TRUE, main = "", xlab = "NH3", ylab = "")
hist(df$SO2, probability = TRUE, main = "", xlab = "SO2", ylab = "")
hist(df$Ozone, probability = TRUE, main = "", xlab = "Ozone", ylab = "")
```

\center{Figure 1: The above graph shows the histogram of values of the six pollutant parameters.}

\newpage

```{=tex}
\flushleft
\textbf{\large A Pair Plot}
```

\flushleft{An important part of exploratory data analysis is to find out which parameters are correlated, positive or negative. This helps one decide on which parameters to include in their prediction model. Since there are 10,600 data points in our data and also since all the stations are present in Delhi, without the loss of generality, we will look at the pair plot of parameter values for the station Rohini, station code 1430.}\
\
\

```{r, echo=FALSE,  fig.width=10, fig.height=10}
pairs(~ PM2.5 + PM10 + NO2 + NH3 + SO2  +Ozone, data = df[df$siteName=="Rohini",], pch=20,
      col = rgb(0, 0, 0, 0.15))
```

\center{Figure 2: The above graph shows the pair plot for all the six pollutant parameters.}

\newpage

```{=tex}
\flushleft
\textbf{\large A Correlation Plot}
```
\flushleft{A numerical view to pair plot is a correlation plot. The patterns between pollutant parameters we observed in the pair plot are expressed as a number between -1 and 1 called correlation coefficient. Here, we are not differentiating between positive correlation and negative correlation. We just look if they are correlated be it positive or negative. Hence we plot the absolute value of the correlation coefficient between parameters.}\
\
<!-- \flushleft{As going witht the theme of grey pallete we adjust the correlation between light grey to dark grey, where light grey represent no correlation and dark grey represent complete correlation.}\ -->
<!-- \ -->
<!-- \ -->

```{r, echo=FALSE, fig.width=7, fig.height=4}
cor_df <- abs(round(cor(df[4:9]), 2))
melted_cor <- melt(cor_df)

ggplot(data = melted_cor, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  geom_text(aes(Var2, Var1, label = value), size = 5) +
  scale_fill_gradient2(low = "lightgrey", high = "darkgrey",
                       limit = c(0,1), name="Absolute\nCorrelation\n") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_text(size=12), axis.text.y = element_text(size=12),
        panel.background = element_blank())
```
\center{Figure 3: The above graph shows the correlation plot for all six pollutants in the data.}
\vspace{1.5cm}

```{r, echo=FALSE, fig.width=6, fig.height=4, fig.align = 'left'}
x = df$PM2.5
y = df$PM10
plot(df$PM2.5, df$PM10, pch=20,col = rgb(0, 0, 0, 0.15), main = "", xlab = "PM2.5", ylab = "PM10")
abline(lm(y ~ x, data = df), col = "red", lw = 2, lty=6)
```
\center{Figure 4: Scatter plot of highest correlated parameters PM2.5 and PM10.}
\newpage

```{=tex}
\flushleft
\textbf{\large A Time-Series Plot}
```

\flushleft{Since we have a time-stamped data, one of the basic analysis in time-series is to find the trends across time. For the sake of simplicity we will show time series plots for only one parameter i.e. PM2.5 for all the stations. The y-axis has been adjusted to same scale for ease of analysis.}\
\
\

```{r, echo=FALSE, fig.width=15, fig.height=12}
par(mar=c(2,2,2,2))
par(mfrow=c(5,2))
for (i in unique(df$siteName)) {
  plot(df[df$siteName==i,]$Date, df[df$siteName==i,]$PM2.5, type = "l",
  main = i, xlab = "", ylab = "", ylim = c(0,700))
}
```

\center{Figure 5: The above graph shows the time series plot of PM2.5 parameter for all 10 stations in the data.}

\newpage

```{=tex}
\flushleft
\textbf{\large Another Time-Series Plot}
```

\flushleft{As stated previously, our main objective is to look for trends. This plot is same as the previous plot but added with educated guess of ticks where there is a trend. A regular jump in values of pollutant parameter PM2.5 between the month of October and February every year. The colued region is from Oct 15 to Feb 15 every year. This trend can be observed uniformly at all ten stations.}\
\
\

```{r, echo=FALSE, fig.width=15, fig.height=12}
par(mar=c(2,2,2,2))
par(mfrow=c(5,2))
for (i in unique(df$siteName)) {
  x0 = df[df$siteName==i,]$Date
  y0 = df[df$siteName==i,]$PM2.5
  x1 = x0+1
  y1 = y0[2:length(y0)]
  cond1 = ((x0>as.Date("2018-10-15") & x0<as.Date("2019-02-15")))
  cond2 = ((x0>as.Date("2019-10-15") & x0<as.Date("2020-02-15")))
  cond3 = ((x0>as.Date("2020-10-15")))
  colour1 <- ifelse((cond1 | cond2 | cond3),"red", "black")
  plot(x0, y0, type="n", main = i, xlab = "", ylab = "", ylim = c(0,700), lwd.ticks=2)
  segments(x0 = x0, y0 = y0, x1 = x1, y1 = y1, col = colour1)
  
  axis(1, at = as.Date("2018-10-15"), labels = "Oct", lwd.ticks=2)
  axis(1, at = as.Date("2019-10-15"), labels = "Oct", lwd.ticks=2)
  axis(1, at = as.Date("2020-10-15"), labels = "Oct", lwd.ticks=2)
  
  axis(1, at = as.Date("2019-02-15"), labels = "Feb", lwd.ticks=2)
  axis(1, at = as.Date("2020-02-15"), labels = "Feb", lwd.ticks=2)
}
```

\flushleft{Figure 6: The above graph shows the time series plot of PM2.5 parameter for all 10 stations in the data highlighted in red with the interval where there is a jump in PM2.5 values.}

\newpage

## Results

* From Figure 1, we see that PM2.5 resembles exponential and PM10 and Ozone resembles as left-skewed normal.


* From Figure 2, most of the parameters seem to have slightly correlation with each other, while we suspect PM2.5 and PM10 to be highly correlated as all the particles in PM2.5 are included in PM10.


* From Figure 3, we see that in fact our guess from Figure 2 is correct. PM2.5 and PM10 do actually have a strong correlation of 0.89 while other are slightly correlated


* From Figure 4, we see that both PM2.5 and PM10 have linear relation.


* From Figure 5, we see that there are subtle jumps in value of PM2.5 parameter across all the stations and repeats at equal interval.


* From Figure 6, we see that in fact the jumps are between the moth of October and February.

\vspace{2cm}

## Conclusion
Recalling our problem statement,

*In this report we try to discover patterns and make inferences about the pollution level in stations in and around Delhi. We also look at the relation between the pollutant parameters and try to find which parameter are related to each other and see if there is a pattern across all ten stations.*

We do conclude that:

**Yes**, the pollutants are related.

**Yes**, a pattern can be observed across all the stations.

